### 2012-03-25 exam4/2.tex
problems 1&2 with solns

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{tikz} 
\usepackage{qtree} 
\usepackage{./exam}
\usepackage{type1cm} % We are using scalable fonts. Really.
\usepackage{alltt,times,comment,amsmath}
\usepackage{alltt}
\usepackage{ct} % tight CM font for \tt
\usepackage{xcolor}
\newexercise{exercise}{Exercise}[subsection]

@@

\ifrubric
\newenvironment{solution}{\color{Red}}{}
\else
\excludecomment{solution}
\fi
\newexercise{problem}{Problem}

@@

\emph{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{6cm}
\rule{1cm}{0pt}\begin{tabular}{|c|l|@{/}r|}
\hline
{\bf Problem} & Points & out of \\ \hline
1 & & ??\\ \hline
2 & & ??\\ \hline
3 & & ??\\ \hline
4 & & ??\\ \hline
5 & & ??\\ \hline
%% Extra & & 5 \\ \hline
{\bf Total} & & ?? \\ \hline
%{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
\end{tabular}
\end{minipage}

@@

\begin{problem}\pts{18} % same as on the previous exam :)

@@

\Tree [. {\tt "Cat"} {\tt "Hat"} [. {\tt "Zebra"} {\tt "Rhino"}  ] !{\qbalance} ]

@@

\item Design the {\tt height} method, which computes the height of the
  tree (i.e., the maximum distance from the root to a leaf).  The
  example tree has a height of 3.

@@

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi

\begin{solution}
\begin{alltt}
\rubricpts{? pts}
\end{alltt}
\end{solution}
\end{problem}
%\ifrubric\bigskip\else
\newpage
%\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{10}

Here are data, class, and interface definitions for a list of numbers:
\begin{alltt}
;; An LoN is one of:
;; - (new mt%)
;; - (new cons% Number LoN)
;; and implements:
;; - sum : -> Number
;;   Compute the sum of the numbers in this list.
;; - sqrs : -> LoN
;;   List of squares of each number in this list.

(define-class mt%
  (define (sum) 0)
  (define (sqrs) this))

(define-class cons%
  (fields first rest)
  (define (sum)
    (+ (this . first)
       (this . rest . sum)))
  (define (sqrs)
    (new cons% 
         (sqr (this . first)) ; (sqr x) = x\(\sp{2}\)
         (this . rest . sqrs))))
\end{alltt}
%
Here are some examples:
\begin{verbatim}
(check-expect ((new mt%) . sum) 0)
(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum)  9)
(check-expect ((new mt%) . sqrs) (new mt%))
(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sqrs)
              (new cons% 16 (new cons% 25 (new mt%))))
\end{verbatim}

\newpage
\noindent
Develop the \verb|sum-sqrs| method that computes the sum of the
squares of a list of numbers.  Here is an example:
\begin{verbatim}
(check-expect ((new cons% 4 (new cons% 5 (new mt%))) . sum-sqrs) 
              (+ 16 25))
\end{verbatim}

If you can, define a super class of \verb|mt%| and \verb|cons%| and
lift the definition of \verb|sum-sqrs| to this class.  (We will assume
the appropriate \verb|super| declaration in \verb|mt%| and
\verb|cons%|.)  If you cannot, define methods in both \verb|mt%| and
\verb|cons%| for partial credit.

\begin{solution}
\begin{alltt}
\rubricpts{2 pts% -- class def}
(define-class list%
  \rubricpts{2 pts -- contract and purpose}
  ;; sum-sqrs : -> Number
  ;; compute the sum of the squares of the numbers in this list
  \rubricpts{4 pts -- implementation}
  (define (sum-sqrs)
    ((this . sqrs) . sum)))

\rubricpts{2 pts -- tests}
(check-expect ((new mt%) . sum-sqrs) 0)
\end{alltt}
\end{solution}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{tikz} 
\usepackage[]{xcolor}
\usepackage{qtree} 
\usepackage{./exam}
\usepackage{type1cm} % We are using scalable fonts. Really.
\usepackage{alltt,times,comment,amsmath}
\usepackage{alltt}
\usepackage{ct} % tight CM font for \tt
\newexercise{exercise}{Exercise}[subsection]

@@

\ifrubric
\newenvironment{solution}{\color{red}}{}
\else
\excludecomment{solution}
\fi
\newexercise{problem}{Problem}

@@

\emph{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{6cm}
\rule{1cm}{0pt}\begin{tabular}{|c|l|@{/}r|}
\hline
{\bf Problem} & Points & out of \\ \hline
1 & & 15\\ \hline
2 & & 22\\ \hline
3 & & ??\\ \hline
4 & & ??\\ \hline
5 & & ??\\ \hline
%% Extra & & 5 \\ \hline
{\bf Total} & & ?? \\ \hline
%{\bf Base}  & \multicolumn{2}{l|}{{\bf 56}} \\ \hline
\end{tabular}
\end{minipage}

@@

\begin{problem}\pts{15} % same as on the previous exam :)

@@

\Tree [. {\tt "Bread"} {\tt "Hat"} [. [. {\tt "Zebra"} {\tt "Dog"} {\tt
  "Cat"} ] {\tt "Rhinocerous"}  ] !{\qbalance} ]

@@

\item Design the {\tt height} method, which computes the height of the
  tree (i.e., the maximum distance from the root to a leaf).  The
  example tree has a height of 4.

@@

\ifrubric\else
\newpage
[Here is some more space for the previous problem.]
\newpage
[Here is some more space for the previous problem.]
\newpage
\fi

\begin{solution}
\begin{verbatim}
//> 4pts for interfaces
interface Tree {
  //> 1pt for purpose
  // count the height of this tree
  Integer height();
}

interface TwoTree extends Tree {}
interface ThreeTree extends Tree {}
\end{verbatim}
\newpage
\begin{verbatim}
//> 4pts for class definitions
//> 3pts for method definitions
class Leaf implements TwoTree, ThreeTree {
  String val;

  Integer height() { return 1; }
}

class TwoNode implements TwoTree {
  ThreeTree left;
  ThreeTree right;

  Integer height() { return max(this.left.height(), this.right.height()); }
}

class ThreeNode implements ThreeTree {
  TwoTree left;
  TwoTree mid;
  TwoTree right;

  Integer height() { return max(this.left.height(),
                                this.mid.height(),
                                this.right.height()); }
}

Tree r = new Leaf("Rhinocerous");
Tree h = new Leaf("Hat");
Tree b = new Leaf("Bread");
Tree t1 = new ThreeNode(new Leaf("Zebra"),new Leaf("Dog"),new Leaf("Cat"));
Tree t2 = new TwoNode(t1,r);
Tree t3 = new ThreeNode(b,h,t2);

//> 3pts for tests
t3.height -> 4
r.height -> 1
t2.height -> 3
\end{verbatim}
\end{solution}
\end{problem}
%\ifrubric\bigskip\else
\newpage
%\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{22}

Design a \verb|TreeVisitor| interface for 2/3-trees, and implement the
\verb|visit| method in all of your classes from the previous problem.
Make sure you design the interfaces and methods so that a visitor can
return any type of data.

\begin{solution}
\begin{verbatim}
//> 2pts parameterization
//> 2pts methods
interface TreeVisitor<T> {
  T visitLeaf(String s);
  T visitTwoNode(ThreeTree l, ThreeTree r);
  T visitThreeNode(TwoTree l, TwoTree r);
}

//> 2pts overall for getting types right
//> 3pts overall for getting impls right

// in Leaf:
<T> T visit(TreeVisitor<T> v) { 
  return v.visitLeaf(this.val);
}

// in TwoNode:
<T> T visit(TreeVisitor<T> v) { 
  return v.visitTwoNode(this.left, this.right);
}

// in ThreeNode:
<T> T visit(TreeVisitor<T> v) { 
  return v.visitThreeNode(this.left, this.mid, this.right);
}
// no tests required!
\end{verbatim}
\end{solution}

\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
\fi

Design the \verb|CountVisitor| class, which counts the number of
leaves in a 2/3-tree (there are 6 in the given example.)

\begin{solution}
\begin{verbatim}
//> 2pts class def
//> 2pts implementations
class CountVisitor implements TreeVisitor<Integer> {
  Integer visitLeaf(String s) { return 1; }
  Integer visitTwoNode(Tree l, Tree r) { return l.visit(this) + r.visit(this); }
  Integer visitThreeNode(Tree l, Tree m, Tree r) { 
     return l.visit(this) + r.visit(this) + m.visit(this);
  }
}
//> 2pts tests
r.visit(new CountVisitor()) -> 1
t2.visit(new CountVisitor()) -> 4
t3.visit(new CountVisitor()) -> 6
\end{verbatim}
\end{solution}

Design the \verb|LongestVisitor| class, which produces the longest
\verb|String| in a 2/3-tree.  The longest string in the example is
\verb|"Rhinocerous"|.

\begin{solution}
\begin{verbatim}
// 2pts class def
// 3pts methods
class LongestVisitor implements TreeVisitor<String> {
  String visitLeaf(String s) { return s; }
  String visitTwoNode(Tree l, Tree r) { 
     String s1 = l.visit(this);
     String s2 = r.visit(this);
     return (s1.length() > s2.length())?s1:s2;
  }
  String visitThreeNode(Tree l, Tree m, Tree r) { 
     String s1 = l.visit(this);
     String s2 = r.visit(this);
     String s3 = m.visit(this);
     if (s1.length() > s2.length() && s1.length() > s3.length())
       return s1;
     else if (s2.length() > s3.length())
       return s2;
     else
       returns s3;
  }
}
//> 2pts tests
h.visit(new LongestVisitor()) -> "Hat"
t2.visit(new LongestVisitor()) -> "Rhinocerous"
t3.visit(new LongestVisitor()) -> "Rhinocerous"
\end{verbatim}
\end{solution}
